---
title: "London Weather Analysis"
date: "Thursday, 12/15"
author: "[Chenjia Li]"
output: html_notebook
---

```{r}
## Import packages
#rm(list = ls())

#packages
library(imputeTS)
library(zoo)
library(forecast)
library(bayesrules)
library(randomForest)
library(rpart)
library(tree)
library(pROC)
library(mgcv)
library(ISLR)
library(lars)
library(dplyr)
library(tidyverse)
library(faraway)
library(olsrr)
library(caret)
library(astsa)
library(glmnet)
library(leaps)
```


## Curvature data maker
```{r}
#curvature maker
matrix.2ndorder.make<-function(x, only.quad=F){
  x0<-x
  dimn<-dimnames(x)[[2]] #extract the names of the variables
  num.col<-length(x[1,]) # how many columns
  for(i in 1:num.col){
    # if we are doing all 2nd order
    if(!only.quad){
      for(j in i:num.col){
        x0<-cbind(x0,x[,i]*x[,j])
        dimn<-c(dimn,paste(dimn[i],dimn[j],sep=""))
        #create interaction dimnames
      }
    } else {
      #in here only if doing only squared terms
      x0<-cbind(x0,x[,i]*x[,i])
      dimn<-c(dimn,paste(dimn[i],"2",sep="")) # squared dimmension names
    }
  }
  dimnames(x0)[[2]]<-dimn
  x0
}
```


## Autoregression prediction function
```{r}
# ar predicted function
my.ar.predict <- function(arstr,newdat,lag,lag1){
  n1<-length(newdat)
  pred<-NULL
  for(i in 0:(n1-lag-1)){
    v1<-c(1:(lag+i))
    pred<-c(pred,predict(arstr,newdat[v1],n.ahead=lag1)$pred[lag1])
  }
  plot(pred,newdat[-c(1:lag)], xlab = "Predicted Value", ylab = "Data Value")
  cor(pred,newdat[-c(1:lag)])
  pred
}
```


## Deviation calculation function
```{r}
# deviation computation function, both monthly and seasonally
comp.deviation <- function(data.vec){
  
  return(data.vec - mean(data.vec))
  
}
```






## 1. Dataset weather-anomalies-1964-2013
```{r}

weatherData <- read.csv("london_weather.csv")

head(weatherData)

```


#### a) Data Wrangling

```{r}
#https://cran.r-project.org/web/packages/imputeTS/imputeTS.pdf
# Checking each predictor, whether they have NA value
# Conclusion: We have to apply spline on the dataset except date and sunshine
head(weatherData[is.na(weatherData$cloud_cover),])
head(weatherData[is.na(weatherData$sunshine),])
head(weatherData[is.na(weatherData$global_radiation),])
head(weatherData[is.na(weatherData$mean_temp),])
head(weatherData[is.na(weatherData$precipitation),])
head(weatherData[is.na(weatherData$pressure),])
head(weatherData[is.na(weatherData$snow_depth),])


# remedy
approxValues1 <- as.matrix(na_kalman(ts(weatherData$cloud_cover, weatherData$date)))[,1]
approxValues2 <- as.matrix(na_kalman(ts(weatherData$global_radiation, weatherData$date)))[,1]
approxValues3 <- as.matrix(na_kalman(ts(weatherData$mean_temp, weatherData$date)))[,1]
approxValues4 <- as.matrix(na_kalman(ts(weatherData$precipitation, weatherData$date)))[,1]
approxValues5 <- as.matrix(na_kalman(ts(weatherData$pressure, weatherData$date)))[,1]
approxValues6 <- as.matrix(na_kalman(ts(weatherData$snow_depth, weatherData$date)))[,1]



weatherData <- weatherData %>%
  select(-max_temp,-min_temp) %>%
  mutate(cloud_cover = round(approxValues1,1), 
         global_radiation = approxValues2,
         mean_temp = approxValues3,
         precipitation = approxValues4,
         pressure = approxValues5,
         snow_depth = approxValues6)
```


```{r}
#Right now, we have no NA entry in our dataset.
weatherData


head(weatherData[is.na(weatherData$cloud_cover),])
head(weatherData[is.na(weatherData$sunshine),])
head(weatherData[is.na(weatherData$global_radiation),])
head(weatherData[is.na(weatherData$mean_temp),])
head(weatherData[is.na(weatherData$precipitation),])
head(weatherData[is.na(weatherData$pressure),])
head(weatherData[is.na(weatherData$snow_depth),])


weatherData[weatherData$date %in% c(20050912, 20050515, 20051230, 20060109, 20060203, 20060317),]

```


#### Monthly deviation

```{r}
yr <- c(1979:2020)
month <- c("01","02","03","04","05","06","07","08","09","10","11","12")

tempData <- weatherData %>%
  select(date,mean_temp)

Deviation_wd <- NULL

date.char <- substring(as.character(tempData$date),1,6)


for ( i in yr ){
  for ( j in month ){
    tData <- round(comp.deviation(tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(i),j)), date.char)]),2)
    Deviation_wd <- c(Deviation_wd,tData)
  }
}

MonthData <- weatherData %>%
  mutate(mean_temp = Deviation_wd) %>%
  rename(Month_Dev = mean_temp)

MonthData

```



#### Seasonally deviation

```{r}
yr <- c(1979:2019)

tempData2 <- weatherData %>%
  select(date,mean_temp)

season_DevData <- NULL

t1Data <- round(comp.deviation(tempData$mean_temp[date.char == "197901"]),2)
season_DevData <- c(season_DevData,t1Data)

for ( i in yr ){
  season_DevData <- c(season_DevData,round(comp.deviation(c(tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(i),"02")), date.char)],tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(i),"03")), date.char)],tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(i),"04")), date.char)])),2))
  
  season_DevData <- c(season_DevData,round(comp.deviation(c(tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(i),"05")), date.char)],tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(i),"06")), date.char)],tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(i),"07")), date.char)])),2))
  
  season_DevData <- c(season_DevData,round(comp.deviation(c(tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(i),"08")), date.char)],tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(i),"09")), date.char)],tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(i),"10")), date.char)])),2))
  
  season_DevData <- c(season_DevData,round(comp.deviation(c(tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(i),"11")), date.char)],tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(i),"12")), date.char)],tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(i+1),"01")), date.char)])),2))
}

season_DevData <- c(season_DevData,round(comp.deviation(c(tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(2020),"02")), date.char)],tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(2020),"03")), date.char)],tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(2020),"04")), date.char)])),2))
  
season_DevData <- c(season_DevData,round(comp.deviation(c(tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(2020),"05")), date.char)],tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(2020),"06")), date.char)],tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(2020),"07")), date.char)])),2))
  
season_DevData <- c(season_DevData,round(comp.deviation(c(tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(2020),"08")), date.char)],tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(2020),"09")), date.char)],tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(2020),"10")), date.char)])),2))
  
season_DevData <- c(season_DevData,round(comp.deviation(c(tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(2020),"11")), date.char)],tempData$mean_temp[grepl(gsub(" ", "", paste(as.character(2020),"12")), date.char)])),2))
  

SeasonData <- weatherData %>%
  mutate(mean_temp = season_DevData) %>%
  rename(Season_Dev = mean_temp)

SeasonData
```




## 2. Time series on 1)original Dataset 2)Monthly temperature Deviation 3)Seasonal temperature Deviation


#### 1) Monthly Dataset

```{r}

#Transform data frame to time series class
ts.o <- ts(weatherData$mean_temp, start = 1, end = nrow(weatherData), frequency = 1)
#Visualize ts
plot(ts.o)
#Length of data
dlen<- length(ts.o)
#length of model data
lm.o <- length(ts.o) - 3500
#Build model autoregression
ar.o <- ar(ts.o[1:lm.o])
#lag selected by autoregrssion model
lag.o<- ar.o$order


dataVal <- weatherData$mean_temp[-c(1:lm.o)]
dataVal <- dataVal[-c(1:lag.o)]

#predicton based on different lag
pred1.o <- my.ar.predict(ar.o,ts.o[-c(1:lm.o)],lag.o,1)
pred2.o <- my.ar.predict(ar.o,ts.o[-c(1:lm.o)],lag.o,2)
pred3.o <- my.ar.predict(ar.o,ts.o[-c(1:lm.o)],lag.o,3)
pred4.o <- my.ar.predict(ar.o,ts.o[-c(1:lm.o)],lag.o,4)
pred5.o <- my.ar.predict(ar.o,ts.o[-c(1:lm.o)],lag.o,5)
pred6.o <- my.ar.predict(ar.o,ts.o[-c(1:lm.o)],lag.o,6)
pred7.o <- my.ar.predict(ar.o,ts.o[-c(1:lm.o)],lag.o,7)



mae1.o <- median(abs(pred1.o-dataVal))
mae2.o <- median(abs(pred2.o-dataVal))
mae3.o <- median(abs(pred3.o-dataVal))
mae4.o <- median(abs(pred4.o-dataVal))
mae5.o <- median(abs(pred5.o-dataVal))
mae6.o <- median(abs(pred6.o-dataVal))
mae7.o <- median(abs(pred7.o-dataVal))
mae1.o
mae2.o
mae3.o
mae4.o
mae5.o
mae6.o
mae7.o

plot(c(1:7), c(mae1.o,mae2.o,mae3.o,mae4.o,mae5.o,mae6.o,mae7.o), xlab = "lag", ylab = "MAE")

```



#### 2) Monthly Dataset

```{r}

#Transform data frame to time series class
ts.m <- ts(MonthData$Month_Dev, start = 1, end = nrow(MonthData), frequency = 1)
#Visualize ts
plot(ts.m)
#Length of data
dlen.m<- length(ts.m)
#length of model data
lm.m <- length(ts.m) - 3500
#Build model autoregression
ar.m <- ar(ts.m[1:lm.m])
#lag selected by autoregrssion model
lag.m<- ar.m$order


dataVal.m <- MonthData$Month_Dev[-c(1:lm.m)]
dataVal.m <- dataVal.m[-c(1:lag.m)]

#predicton based on different lag
pred1.m <- my.ar.predict(ar.m,ts.m[-c(1:lm.m)],lag.m,1)
pred2.m <- my.ar.predict(ar.m,ts.m[-c(1:lm.m)],lag.m,2)
pred3.m <- my.ar.predict(ar.m,ts.m[-c(1:lm.m)],lag.m,3)
pred4.m <- my.ar.predict(ar.m,ts.m[-c(1:lm.m)],lag.m,4)
pred5.m <- my.ar.predict(ar.m,ts.m[-c(1:lm.m)],lag.m,5)
pred6.m <- my.ar.predict(ar.m,ts.m[-c(1:lm.m)],lag.m,6)
pred7.m <- my.ar.predict(ar.m,ts.m[-c(1:lm.m)],lag.m,7)



mae1.m <- median(abs(pred1.m-dataVal.m))
mae2.m <- median(abs(pred2.m-dataVal.m))
mae3.m <- median(abs(pred3.m-dataVal.m))
mae4.m <- median(abs(pred4.m-dataVal.m))
mae5.m <- median(abs(pred5.m-dataVal.m))
mae6.m <- median(abs(pred6.m-dataVal.m))
mae7.m <- median(abs(pred7.m-dataVal.m))
mae1.m
mae2.m
mae3.m
mae4.m
mae5.m
mae6.m
mae7.m

plot(c(1:7), c(mae1.m,mae2.m,mae3.m,mae4.m,mae5.m,mae6.m,mae7.m), xlab = "lag", ylab = "MAE")


```







#### 3) Seasonal Dataset

```{r}

#Transform data frame to time series class
ts.s <- ts(SeasonData$Season_Dev, start = 1, end = nrow(SeasonData), frequency = 1)
#Visualize ts
plot(ts.s)
#Length of data
dlen.s<- length(ts.s)
#length of model data
lm.s <- length(ts.s) - 3500
#Build model autoregression
ar.s <- ar(ts.s[1:lm.s])
#lag selected by autoregrssion model
lag.s<- ar.s$order


dataVal.s <- SeasonData$Season_Dev[-c(1:lm.s)]
dataVal.s <- dataVal.s[-c(1:lag.s)]

#prediction based on different lag
pred1.s <- my.ar.predict(ar.s,ts.s[-c(1:lm.s)],lag.s,1)
pred2.s <- my.ar.predict(ar.s,ts.s[-c(1:lm.s)],lag.s,2)
pred3.s <- my.ar.predict(ar.s,ts.s[-c(1:lm.s)],lag.s,3)
pred4.s <- my.ar.predict(ar.s,ts.s[-c(1:lm.s)],lag.s,4)
pred5.s <- my.ar.predict(ar.s,ts.s[-c(1:lm.s)],lag.s,5)
pred6.s <- my.ar.predict(ar.s,ts.s[-c(1:lm.s)],lag.s,6)
pred7.s <- my.ar.predict(ar.s,ts.s[-c(1:lm.s)],lag.s,7)



mae1.s <- median(abs(pred1.s-dataVal.s))
mae2.s <- median(abs(pred2.s-dataVal.s))
mae3.s <- median(abs(pred3.s-dataVal.s))
mae4.s <- median(abs(pred4.s-dataVal.s))
mae5.s <- median(abs(pred5.s-dataVal.s))
mae6.s <- median(abs(pred6.s-dataVal.s))
mae7.s <- median(abs(pred7.s-dataVal.s))
mae1.s
mae2.s
mae3.s
mae4.s
mae5.s
mae6.s
mae7.s

plot(c(1:7), c(mae1.s,mae2.s,mae3.s,mae4.s,mae5.s,mae6.s,mae7.s), xlab = "lag", ylab = "MAE")


```

# Overall Deviation
```{r}

#Transform data frame to time series class
dev4 <- comp.deviation(weatherData$mean_temp)
data_overall <- weatherData %>%
  mutate(dev_overall = dev4) %>%
  select(-mean_temp)

#sample lag (= 1)
temp_lag <- data_overall$dev_overall
temp_lag <- c(temp_lag, NA)
temp_lag <- temp_lag[-1]

ts.4 <- ts(dev4, start = 1, end = nrow(weatherData), frequency = 1)
#Visualize ts
plot(ts.4)
#Length of data
dlen.4<- length(ts.4)
#length of model data
lm.4 <- length(ts.4) - 3500
#Build model autoregression
ar.4 <- ar(ts.4[1:lm.4])
#lag selected by autoregrssion model
lag.4<- ar.4$order


dataVal.4 <- dev4[-c(1:lm.4)]
dataVal.4 <- dev4[-c(1:lag.4)]

#prediction based on different lag
pred1.4 <- my.ar.predict(ar.4,ts.4[-c(1:lm.4)],lag.4,1)
pred2.4 <- my.ar.predict(ar.4,ts.4[-c(1:lm.4)],lag.4,2)
pred3.4 <- my.ar.predict(ar.4,ts.4[-c(1:lm.4)],lag.4,3)
pred4.4 <- my.ar.predict(ar.4,ts.4[-c(1:lm.4)],lag.4,4)
pred5.4 <- my.ar.predict(ar.4,ts.4[-c(1:lm.4)],lag.4,5)
pred6.4 <- my.ar.predict(ar.4,ts.4[-c(1:lm.4)],lag.4,6)
pred7.4 <- my.ar.predict(ar.4,ts.4[-c(1:lm.4)],lag.4,7)

mae1.4 <- median(abs(pred1.4-dataVal.4))
mae2.4 <- median(abs(pred2.4-dataVal.4))
mae3.4 <- median(abs(pred3.4-dataVal.4))
mae4.4 <- median(abs(pred4.4-dataVal.4))
mae5.4 <- median(abs(pred5.4-dataVal.4))
mae6.4 <- median(abs(pred6.4-dataVal.4))
mae7.4 <- median(abs(pred7.4-dataVal.4))
mae1.4
mae2.4
mae3.4
mae4.4
mae5.4
mae6.4
mae7.4

plot(c(1:7), c(mae1.4,mae2.4,mae3.4,mae4.4,mae5.4,mae6.4,mae7.4), xlab = "lag", ylab = "MAE")

```

```{r}
data_overall
```



## 3. Arima Model Forecasting

ARIMA (AutoRegressive Integrated Moving Average) and AR (AutoRegressive) are both time series forecasting methods. The main difference between the two is that ARIMA includes the additional concept of integration, which accounts for the fact that the time series may be non-stationary. In other words, ARIMA can be used to model time series that have a trend or seasonal component, while AR can only be used for stationary time series.In this reason, we would like to see whether Arima works better than AR.

```{r}
arima.len <- lm.m+lag.m
arima.m <- auto.arima(ts.m[1:arima.len])

predict(arima.m,newdata=ts.m[-c(1:arima.len)])

pred_arima.m <- forecast(arima.m,3460)$mean
pred_arima.m <- as.matrix(pred_arima.m)[,1]
median(abs(dataVal.m - pred_arima.m))


arima.len1 <- lm.s+lag.s
arima.s <- auto.arima(ts.s[1:arima.len1])

predict(arima.s,newdata=ts.s[-c(1:arima.len1)])

pred_arima.s <- forecast(arima.s,3494)$mean
pred_arima.s <- as.matrix(pred_arima.s)[,1]
median(abs(dataVal.s - pred_arima.s))
```



## 4. Model Selection and Regression Model

#### a) Scatter Plot
The scatter Plot shows that we have to deal with the multicollinearity and overfitting problem by Larsso
```{r}
pairs(weatherData)
```

#b) We shift the temperature to the day before, which means we used today's predictors to predict tomorrow mean temperature
```{r}
newTemp <- weatherData$mean_temp
newTemp <- c(newTemp, NA)
newTemp <- newTemp[-1]
newData <- weatherData %>%
  select(-date) %>%
  mutate(mean_temp = newTemp)
newData <- newData[-nrow(newData),]
nrow(newData)
newData
```


#c) Split the original dataset into trainSet and Test set
```{r}
new.len <- nrow(newData) - 3500
matData <- newData
trainData <- matData[c(1:new.len),]
testData <- matData[-c(1:new.len),]
```

```{r}
lmod <- lm(mean_temp ~. ,data = trainData)
summary(lmod)
```

It is possible for a linear regression model to have all significant regressors but still not be the best model for a given data set. This can happen if the model is overfitted to the training data, meaning that it has a high degree of flexibility and can fit the training data very well, but does not generalize well to new data.

So we would like to implement model selection and other criteria to find the best model for prediction.




## d) Use train set for model building (2nd order data and 2nd order with interaction term data). Plot the larsso plot and the plot of degree of freedom against Cp. We need to choose the best Cp from the output.

```{r}
y.train <- as.matrix(trainData[,4])
x.train <- as.matrix(trainData[,-4])

#Only second order 
weather.mat2nd.t <- matrix.2ndorder.make(x.train, only.quad=T)
dum.lars.t <- lars(weather.mat2nd.t,y.train)
plot(dum.lars.t)
dum.t <- dum.lars.t
plot(dum.t$df,dum.t$Cp,log="y")
lines(dum.t$df,dum.t$df)


#include interaction term
weather.mat2nd.f <- matrix.2ndorder.make(x.train,only.quad=F)
dum.lars.f <- lars(weather.mat2nd.f, y.train)
plot(dum.lars.f)
dum.f<-dum.lars.f
plot(dum.f$df,dum.f$Cp,log="y")
lines(dum.f$df,dum.f$df)
```



#### e)Find the best Cp
```{r}
#From the above graph
#for 2nd
I1 <- dum.t$df==12
dum.t$Cp[I1]

I1<-dum.t$df==13
dum.t$Cp[I1]


#for interaction
#test df > 25
I1<-dum.f$df==25
dum.f$Cp[I1]

I1 <- dum.f$df==26
dum.f$Cp[I1]

I1<-dum.f$df==27
dum.f$Cp[I1]

I1<-dum.f$df==28
dum.f$Cp[I1]

#Chose Cp = 59
coef(dum.f)[59,]
```



## f)Use the best Cp to predict the response value of test set and find its Median Absolute value
```{r}
y.test <- as.matrix(testData[,4])
x.test <- as.matrix(testData[,-4])
Snd.test <- matrix.2ndorder.make(x.test,only.quad=T)
diff <- mean(y.test[,1]) - mean(Snd.test%*%coef(dum.t)[14,])
mae <- median(abs(y.test[,1] - Snd.test%*%coef(dum.t)[14,] - diff))
mae


Snd.test <- matrix.2ndorder.make(x.test,only.quad=F)
diff <- mean(y.test[,1]) - mean(Snd.test%*%coef(dum.f)[59,])
mae <- median(abs(y.test[,1] - Snd.test%*%coef(dum.f)[59,] - diff))
mae
```







